Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-06-20 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: add response to comment #4.

We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam, with a random Young's modulus \SCComment{$m(x)$ of length $L$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.  The beam position $x$ is discretized at $N=31$ equidistant points, and 
the vertical deflection $u(x)$ when the beam is subject to a given external load  $g(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = g(x),$$
%where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The  
with the following boundary conditions:
$$ \left.u(x)\right|_{x=0} = 0, \qquad 
\left.\frac{\partial u(x)}{\partial x}\right|_{x=0} = 0, \qquad
\left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{g}$, where $\hat{u}$ and $\hat{g} \in \mathbb{R}^N$. }
%approximates $u(x)$ and $m(x)$ at the nodes, respectively. 
%The forward model $F:\mathbb{R}^3 \rightarrow \mathbb{R}^N $ maps $m = \left[\begin{matrix}m_1, m_2, m_3 \end{matrix}\right]^T$ to $\hat{u}$, while $Q : \mathbb{R}^N \rightarrow \mathbb{R}^2 $ selects the estimated solution at the 10th and 25th nodes, i.e., $ Q(\hat{u}) := \left[\begin{matrix} \hat{u}_{10}, \hat{u}_{25} \end{matrix} \right]^T$. The objective is to determine the distributions of $Q(F(m))$. 
QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.

@@

\SCComment{Response to comment \#2: In this paper, we have focused predominantly on employing QMC sampling in estimating the expectations of functions. However, it can be extended to other important mathematical problems such as function approximation, simulation of Markov chains,  density estimation, and many MC-based randomized algorithms. For instance, Array-RQMC has proven invaluable for maintaining and sorting arrays of Markov chains \cite{l2018sorting,puchhammer2021variance}.  
The SSJ library~\cite{l2002ssj} is a comprehensive tool kit that includes a wide range of QMC point sets, sequences, and randomizations in the ``hups'' and ``mcqmctools'' packages. It will be of great interest to the QMCPy community to make connections to these and other well established software in order to extend its applicability to an increasingly wider scope of  problems in computational mathematics and statistics.
}

\SCComment{Perhaps we could respond to referee's comment \#4 here.}

>>>>

We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam, with a random Young's modulus \SCComment{$m(x)$ of length $L$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.  The beam position $x$ is discretized at $N=31$ equidistant points, and 
the vertical deflection $u(x)$ when the beam is subject to a given external load  $g(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = g(x),$$
%where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The  
with the following boundary conditions:
$$ \left.u(x)\right|_{x=0} = 0, \qquad 
\left.\frac{\partial u(x)}{\partial x}\right|_{x=0} = 0, \qquad
\left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{g}$, where $\hat{u}$ and $\hat{g} \in \mathbb{R}^N$. }
%approximates $u(x)$ and $m(x)$ at the nodes, respectively. 
%The forward model $F:\mathbb{R}^3 \rightarrow \mathbb{R}^N $ maps $m = \left[\begin{matrix}m_1, m_2, m_3 \end{matrix}\right]^T$ to $\hat{u}$, while $Q : \mathbb{R}^N \rightarrow \mathbb{R}^2 $ selects the estimated solution at the 10th and 25th nodes, i.e., $ Q(\hat{u}) := \left[\begin{matrix} \hat{u}_{10}, \hat{u}_{25} \end{matrix} \right]^T$. The objective is to determine the distributions of $Q(F(m))$. 
QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation. % the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.

@@

\SCComment{Response to comment \#2: In this paper, we have focused predominantly on employing QMC sampling in estimating the expectations of functions. However, it can be extended to other important mathematical problems such as function approximation, simulation of Markov chains,  density estimation, and many MC-based randomized algorithms. For instance, Array-RQMC has proven invaluable for maintaining and sorting arrays of Markov chains \cite{l2018sorting,puchhammer2021variance}.  
The SSJ library~\cite{l2002ssj} is a comprehensive tool kit that includes a wide range of QMC point sets, sequences, and randomizations in the ``hups'' and ``mcqmctools'' packages. It will be of great interest to the QMCPy community to make connections to these and other well established software in order to become applicable to an increasingly wider scope of problems and applications of computational mathematics and statistics.
}

\SCComment{Response to comment \#4:  Python is known for its versatility, but it can be slower than highly optimized implementations in languages like Java or C++.  For QMCPy developers, it would be beneficial to investigate such underlying interfaces or consider other code acceleration techniques by leveraging Numba, Joblib, or Ray in the future.
}

\SCComment{Question: Do we have only one referee's report for this paper?}
